---
title: "Index"
author: "Min Heo"
date: "11/14/2021"
output: html_document
---

## Introduction
Our group chose to focus on the impact of mental health on people in a variety of places (such as
the workforce, school, the tech industry, etc.). Mental health is an important factor when
considering the well-being of people which is why our group chose to focus on this. Mental health
can take on many forms but our group chose to focus specifally on commonly known health issues
such as depression and anxiety. Our chosen dataset contains data that takes in the gender, where
they work (and the benefits) as well as treatment (such as care programs or mental health
specialists) while also surveying for the consequences of these mental health issues.

## Summary Information
```{r}
survey <- read.csv("~/INFO201/final-project-MinHeo1/survey.csv")

summary(survey)
library(dplyr)
filtered_data <- filter(survey,Age > 0, Age < 100)
mean_age <- mean(survey$Age)
mean(survey$Age, na.rm = TRUE)

median_age <- median(survey$Age, na.rm = TRUE)

max_age <- max(survey$Age, na.rn = TRUE)

min_age <- min(survey$Age, na.rn = TRUE)

family_history<- survey$family_history

no_family_history_total <- sum(stringr::str_detect(family_history, "No"), na.rm = TRUE)

summary_info <- list(mean_age, median_age, max_age, min_age, no_family_history_total)
```
Based off of the fixed dataset (configuring it so that it was consistent while narrowing down the
scope to select states in the United States) we found that mental health is prominent in all
individuals no matter the state that they are in. Although there were some inconsistencies (such
as some reporting weird ages), the assumptions that could be made are clear. The median age is
around 31, which makes sense when considering that is the prime of their careers (i.e., around
5-10 years into their job, working for a promotion, worrying about salary or family, etc.). We
also calculated the family history, and found that some did have a family history of mental
health although the data is inconsistent (seemingly similar rates of mental health issues among
those with family history and those without) which meant assumptions couldn't be made (or at
least good assumptions). Overall, mental health is an issue that cannot be ignored. It is
prevalent in all states shown and shows similar rates among all states (especially among middle
aged individuals).The mean, min, and max age had unusual values of 79428148.3, -1726, and 9.9 x
10^10. These outliers demonstrate the messiness of this dataset because it was self-reported.
These were left in intentionally to show the bias within this datset (i.e., people may not have
taken the survey seriously). The exuberant data shows that mental health could be trivialized and
most would not report correctly, demonstrating the limitations of a self-reported datset
regarding mental health.

## Aggregate Table Script
```{r}
library(tidytext)
library(dplyr)
library(ggplot2)
rm(list = ls())
df <- read.csv("~/INFO201/final-project-MinHeo1/survey.csv")
df <- df %>% filter(grepl('2014', df$Timestamp))
df$Gender <- toupper(df$Gender)
df$Gender[grep("\\bM\\b",df$Gender)] <- 'MALE'
df$Gender[grep("\\bF\\b",df$Gender)] <- 'FEMALE'
df$Gender[grep("\\bWOMAN\\b",df$Gender)] <- 'FEMALE'
keepGender <- c("MALE", "FEMALE")
df <- df[df$Gender %in% keepGender, ]
df <- 
  df %>%
  group_by(Country) %>%
  filter(family_history == "No") %>%
  filter(Country == "United States") %>% 
  subset(select = c(Timestamp, Age, Gender, Country, state, family_history, no_employees, mental_vs_physical))
mostCommon <-
  df %>%
  count(state, sort = TRUE)
keepState <- as.vector(unlist(mostCommon$state))[1:10]
df <- df[df$state %in% keepState, ]
```
The table is sorted based off of the messiness of the dataset. The data was self-reported so
there were differing values entered (such as some putting "M" for male while others put "male")
which was then recalculated by the table. The table also shows the mental health affects of those
in the United States (and select states within it). This was to narrow down the scope of the
dataset as well as allow more pertinent observations to be made. We found that women were
slightly more likely to report suffering from mental health (although since this is
self-reported, men could have underreported based off of stigma) compared to men. We also found
that bigger states (and bigger businesses) tended to have more employees report suffering from
mental health issues compared to smaller states and businesses.

## Chart 1
```{r}
ggplot(df, aes(x = no_employees)) +
  geom_bar(aes(fill = state)) + facet_grid(. ~ state) + 
  theme(axis.text.x=element_text(angle=90, hjust=1)) + labs(title="Mental Health issues Count by Number of Employees", y="Number of Employees", x="Mental Health issues count")
```
The purpose of this chart is to show the number of employees who experience mental health issues
at businesses with differing numbers of employees (in different states). Based off of the chart,
we can see that on average, the more employees a business has the more likely they are to display
mental health issues (although this could be because it is a bigger sample size). Also, "bigger"
states such as California, Texas, and Washington have higher rates of mental illness than
"smaller" states such as Indiana, Illinois, and Maryland. Another observation that could be made,
though, is that more liberal states such as California and Washington are more likely to report
their issues than other states, although this is just an assumption.

## Chart 2
```{r}
library(tidyverse)
library(dplyr)
library(ggplot2)
survey_df <- read.csv("~/INFO201/final-project-MinHeo1/survey.csv") 
state_df <- survey_df %>%
  filter(Country == "United States") %>%
  group_by(state) %>%
filter(state=="CA" || state=="NY" || state=="WA" || state=="TX" || state=="OR" || state=="IN" || state=="MA" || state=="OH" || state=="TN")
  ordered_no_employees <- factor(state_df$no_employees, levels=c("1-5", "6-25", "26-100", "100-500", "500-1000", "More than 1000"))
ggplot(state_df, aes(x=state, y=ordered_no_employees)) + geom_point(aes(col=coworkers)) + labs(title="Employee amount by Willingness to Discuss Mental Health per State", y="Number of Employees", x="State", color="Willingness")
```
The purpose of this chart was to show the willingness of businesses (with differing number of
employees) to talk about mental health with other coworkers. The graph demonstrates that the more
employees they have the more unwilling they seem to want to talk about mental health with their
coworkers (with Texas the only state willing to talk). Although the graph doesn't have any
concrete assumptions, an assumption that can be made is that the willingness to talk about mental
health is still seemingly taboo. Only a few businesses are willing to talk about mental health
with the majority of them reluctant or against talking about mental health.

## Chart 3
```{r}
my_df <- df[-c(94), ]
View(my_df)

state_group <- my_df %>%
  group_by(state)
View(state_group)
ggplot(my_df, aes(x = state, y = Age, color = state)) +
  geom_violin(size = 1) + labs(title = "The age distribution of different states", x="States", y = "Age distribution") + theme(plot.title = element_text(hjust = 0.5))

```
The purpose of this chart is to show the distribution of mental health issues among different
ages. As seen the bulk of mental health issues occurred around the ages 20-40 which is when most
people are in their "peak" of their careers. Meaning that they have to worry about promotions,
salaries, family, etc. Thus, the chart isn't surprising in any way, although there are some
outlirs (such as Ohio having someone report around the age of 5). Most states don't seem to have
mental health issues after the age of around 45, although this could in part due to the older
generation not reporting (or unwilling to report, because again of the stigma against mental
health). Another interesting observation is that the ages is consistent among all shown states
which shows that the rate (and age) of mental health is similar in states and that mental health
can impact anybody anywhere.
