---
title: "idk analysis"
author: "Min Heo"
date: "5/18/2021"
output: html_document
---

```{r setup, include=FALSE, warning=FALSE}
library("knitr")
source("analysis.R")
```
## Introduction
Incarceration rates are a problem in the United States particularly because of the inherent racism
(whether intentional or not) that is part of them. There are higher rates of incarceration among
the African-American population when compared to other races (especially white people) which
raises the issue of why this is the case. Racism is all too apparent in today's age which and it
carries over into the prison system. Because of the inherent issues within the prison I chose to
analyze the population of male non-white races (such as black, Latin, and Asians) to determine
whether the bias was true or not.

## Summary Information
`mean_blacks_jailed_urban` `mean_lationx_jailed` `highest_blacks_prison` `highest_asians_prison` `blacks_jailed_2018`

From the information calculated with the values I found that the number of African Americans
jailed in 2010 was 283896 while in 2018 it was 247547. This was actually a decline but not by much
which shows the sheer amount of African Americans that are detained. The most amount of African
Americans imprisoned was in New York while the highest number of Asian's in prison was in urban
areas (compared to rural). This shows the distribution of prison systems showing that bigger
cities with bigger populations and urban areas are more likely to jail people compared to smaller
places (which goes against the trend that liberal states are less likely to jail people). The mean
number of latinx in jail was 32.59487 while the mean for African Americans was 0.02531465.
Surprisingly, the number of latinx in prison is substantionally higher than African Americans
although this could be due to the sheer amount of them in the prison systems. Overall though, we
can see that there is implicit (or explicit) bias within the prison system. Especially for latinx
and the African American community. Showing that change is necessary in order to get rid of some
of the bias.
## Map
```{r map, echo=FALSE}
black_prisoned_map 
```
The map shows us that the majority of shooting incidents in 2018 took place closer to the
south-eastern side of the United States rather than the North-western portion of the states. It
also shows us that the specific incidents with the most amount of deaths also take place in the
south-eastern states. Researchers might want to look at why these states are so many incidents as
well as having large amount of deaths in single cases. It could do a lot of good if they were able
to determine the root cause of it. 

## Plot
This plot was made to answer the question of what is the relationship between the amount of people
and injured or killed from the shootings that occurred in 2018 for each month in that year. I
chose to do a line chart that has months for it's x-axis and amount of deaths or injuries for it's
y-axis. It was the best choice because a line chart clearly demonstrates relationships between two
different variables by it's shape. It could also reveal further insights into the time progression
and how it can potentially change data.  

```{r plot, echo=FALSE, fig.width = 9, fig.asp = .62}
year_plot
```

The first insight I saw was that the number of injuries far exceeds that of number of deaths for
each month. This could mean a variety of things such as health care being able to help those
injured and avoid death. However, it is still a tragedy that many people had to suffer because of
these shootings. The second insight that I saw was that the amount of injuries increased for each
month after January before peaking in June and gradually decreasing over the months. While the
deaths don't have such a correlation, the peak is extreme enough from the amount of injured people
to be worth mentoning.  

```{r plot, echo=FALSE, fig.width = 9, fig.asp = .62}
bar_chart_jail
```

The first insight I saw was that the number of injuries far exceeds that of number of deaths for
each month. This could mean a variety of things such as health care being able to help those
injured and avoid death. However, it is still a tragedy that many people had to suffer because of
these shootings. The second insight that I saw was that the amount of injuries increased for each
month after January before peaking in June and gradually decreasing over the months. While the
deaths don't have such a correlation, the peak is extreme enough from the amount of injured people
to be worth mentoning.  
